Excerno analysis

### 2022-08-18 paper_analysis.Rmd
Added files for the analysis with real samples

>>>>

# Using real data

We are going use 6 files from this study ...

## Preprocessing

Let's start by listing the 6 VCF files from the `supplied_results` folder:

```{r}
library(vcfR)
library(R.utils)
library(tidyverse)
vcf_files <- list.files(path="supplied_results",pattern = "vcf", full.names = TRUE)
```

We will be subsetting each of those files to the variants for which we have confident information in the DNA. We can assure that by requiring `GQ >30` (Genotype Quality) on the exome side

```{r}
for (i in 1:length(vcf_files)) {
  name <- vcf_files[i]
  vcf <- read.vcfR(name)
  filter_name <- str_replace_all(name,"vcf","high.vcf.gz")

  gq <-  extract.gt(vcf, element="GQ", as.numeric=TRUE)
  write.vcf(vcf[which(gq[,1]>30),], file=filter_name)
  gunzip(filter_name)
}
```

## Running excerno

We ran excerno on the 6 vcf files and use the NMF option

```{r}
library(MutationalPatterns)
library(BSgenome)
library(excerno)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg38"
library(ref_genome, character.only = TRUE)

vcf_files <- list.files(path="supplied_results",pattern = "high.vcf", full.names = TRUE)
excerno_vcf(vcf_files,method="nmf",2)
```

## Calculating performance metrics

We first create a function that allows to collect information on a filtered vcf file

```{r}
get_metrics <- function (file, sample_name) {
  vcf <- read.vcfR(file)
  gt <- extract.gt(vcf)

  # Calculate the errors in the true variants  
  true_variants <- vcf[which(gt[,1]!="0/0"),]
  all_true <- as.numeric(dim(true_variants)[1])
  true_classify <- getFIX(true_variants)[,7]
  fn <- sum(is.na(true_classify))
  tp <- all_true -fn

  # Calculate the errors in the false variants
  false_variants <- vcf[which(gt[,1]=="0/0"),]
  all_false <- as.numeric(dim(false_variants)[1])
  false_classify <- getFIX(false_variants)[,7]
  tn <- sum(is.na(false_classify))
  fp <- all_false - tn
  sum(is.na(true_classify))

  # Calculate sensitivity and speficifty
  (sens <-  tp /all_true)
  (spec <- tn /all_false)
  (ffpe_pct <- all_false/(all_false+all_true))
  (total <- all_false+all_true)

  tibble(name=sample_name, sens=sens, spec=spec, 
         ffpe.pct = ffpe_pct*100, total=total)
}
```

Now we list the classified files from `excerno`

```{r}
vcf_files <- list.files(pattern = "vcf", full.names = TRUE)
sample_names = c("MLHG10_10", "MSH6_1","MSH6_2", "MSH6_8", "PMS2_10", "PMS2_09")
```

And we generate the table with informaton from the samples

```{r}
1:6 %>%
  map_dfr(~get_metrics(vcf_files[.x], sample_names[.x])) %>%
  arrange(desc(ffpe.pct)) %>%
  mutate_if(is.numeric, format, digits=1)%>%
  kbl() %>%
  kable_styling()
